Hydras \& Co.: Formalized mathematics in Coq\\ for inspiration and entertainment

### 2021-10-15 doc/JFLA2022/paper.tex
new title
pass over genealogy section

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Hydras,  Ordinals \& Co.   \\
  A library in Coq of entertaining formal mathematics}

@@

We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles. In the interest of space, we shall present only the second one in this article.

@@

\gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}. The data types for ordinals are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to placement in different modules. Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral library.
%
Nevertheless, we were able to establish an initial correspondence between both implementations of ordinals through rewriting lemmas. For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine Gaia's corresponding multiplication operation, and were able to import Gaia's proof of associativity of the multiplication almost for free.

\subsection{Project Genealogy, Dependencies, and Organization}

Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions.

\begin{figure}[h]
\centering
{\footnotesize
\begin{tikzpicture}[thick]

@@

Table~\ref{tbl:loc} lists the size of each \Hydras package.

\begin{table}
\centering
\footnotesize
\begin{tabular}{|l|l|l|l|l|}
\hline
\textbf{Package} & \textbf{Repository} & \textbf{Commit} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
\hline
Pocklington & \texttt{pocklington} & & & \\
\hline
Goedel & \texttt{goedel} & & & \\
\hline
Gaia & \texttt{gaia} & & & \\
\hline
Hydra-battles & \texttt{hydra-battles} & & & \\
\hline
Addition-chains & \texttt{hydra-battles} & & & \\
\hline
Gaia-hydras & \texttt{hydra-battles} & & & \\
\hline
\end{tabular}
\caption{Current numbers of lines of code for \Hydras packages.}
\label{tbl:loc}
\end{table}

>>>>

%% Front Matter
%%
% Regular title as in the article class.
%
% \title{Hydras,  Ordinals \& Co.   \\
%   A library in Coq of entertaining formal mathematics}

\title{Hydras \& Co.: Formalized mathematics in Coq\\
 for inspiration and entertainment
}

@@

We chose to build our library and book on two simple themes which allow many variations: computing powers in a monoid, and Kirby and Paris' hydra battles. In the interest of space, we will only present the second theme in this paper.

@@

\gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}. The data types for ordinals are essentially defined the same way in Gaia and \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules. Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral library.
%
Nevertheless, we were able to establish an initial correspondence between both implementations of ordinals through rewriting lemmas. For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine Gaia's corresponding multiplication operation, and were able to import Gaia's proof of associativity of the multiplication almost for free.

\subsection{Package Genealogy, Dependencies, and Organization}

Both due to its prior stages~\cite{PCiota,JFLA2018paper} and the recent integrations described just above, the current \Hydras Coq code has a complex inheritance. Figure~\ref{fig:genealogy} illustrates the relationships between \Hydras packages and show their ancestry from historical Coq contributions. To indicate the scope and relative sizes of packages, Table~\ref{tbl:loc} breaks down lines of code for each \Hydras package, as reported by the \texttt{coqwc} tool. We also name and link to the repository in \community for each package.

\begin{figure}[ht]
\centering
{\footnotesize
\begin{tikzpicture}[thick]

@@

\begin{table}[ht]
\centering
\footnotesize
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Package name} & \textbf{Repository} & \textbf{Package on opam} & \textbf{Version} & \textbf{Spec LOC} & \textbf{Proof LOC}\\
\hline
Pocklington & \href{https://github.com/coq-community/pocklington}{\texttt{pocklington}} & \texttt{coq-pocklington} & \texttt{8.12.0} & & \\
\hline
Goedel & \href{https://github.com/coq-community/goedel}{\texttt{goedel}} & \texttt{coq-goedel} & \texttt{8.13.0} & & \\
\hline
Gaia & \href{https://github.com/coq-community/gaia}{\texttt{gaia}} & \texttt{coq-gaia} & \texttt{1.12} & & \\
\hline
Hydra-battles & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-hydra-battles} & \texttt{0.5} & & \\
\hline
Addition-chains & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-addition-chains} & \texttt{0.5} & & \\
\hline
Gaia-hydras & \href{https://github.com/coq-community/hydra-battles}{\texttt{hydra-battles}} & \texttt{coq-gaia-hydras} & \texttt{0.5} & & \\
\hline
\end{tabular}
\caption{Current numbers of lines of code for \Hydras packages.}
\label{tbl:loc}
\end{table}
